Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome

### 2023-05-23 paper/paper.qmd
more outline for results

```{r load_data}
library(here)
library(tidyverse)

@@

@@

@@

Most important OTUs contributing to model performance (@fig-features)

@@

@@

@@

::: {#fig-features fig-cap="**Feature importance.**
**A)** Feature importance via permutation test.
For each OTU, the order of samples was randomized in the test set 100 times and
the performance was re-calculated to estimate the permutation performance.
An OTU was considered important if the performance decreased when the OTU was
permuted in at least 75% of the models.
OTUs with a greater difference in AUROC (actual performance minus permutation
performance) are more important.
Left: models were trained on the full datasets, with different numbers of
samples available for each severity definition.
Right: models were trained on the intersection of samples with all labels
available for each definition. Note that Attributable and Pragmatic severity are
exactly the same for the intersection dataset.
OTU 120 (*Pseudomonas*) is not shown for IDSA severity on the full dataset nor
on the intersection dataset because it was removed during pre-processing due to
having near-zero variance.
**B)** Log~10~-transformed median relative abundances of the most important OTUs
on the full datasets, grouped by severity (shape).
"}
![](figures/feature-importance.png)
:::

>>>>

```{r load_data, message = FALSE}
library(here)
library(tidyverse)

@@

new dataset.

@@

We first set out to train the best models possible for each severity definition.
Not all samples have labels available for all four severity definitions due to
missing data for some patient lab values and incomplete chart review (TODO table).

However, comparisons across these definitions is not fair when using different
subsets of the data for each definition.
To better compare the model performances across different severity definitions,
we selected the intersection of samples (`r n_cases_int`) that had labels for
all four severity definitions and repeated the model training and evaluation
process.

@@

We performed permutation feature importance to determine which OTUs were the most
important for model performance (@fig-features).

@@

pragmatic vs attributable severity. table?

rough estimate of costs.
current: everyone gets vancomycin.
potential: patients flagged as severe get fidaxomicin. based on NNB, estimate
how much money saved in averting severe outcomes.

@@

Compare to EHR-based models.

Discuss important OTUs. which ones concord with literature, which ones may be new.

Amplicon sequencing is not typically performed for CDI patients, but if there is
clinical value to be gained by implementing OTU-based models, we could make a 
case to routinely sequence and profile the microbial communities of CDI patients.

@@

::: {#fig-features fig-cap="**Feature importance.**
**A)** Feature importance via permutation test.
For each OTU, the order of samples was randomized in the test set 100 times and
the performance was re-calculated to estimate the permutation performance.
An OTU was considered important if the performance decreased when the OTU was
permuted in at least 75% of the models.
OTUs with a greater difference in AUROC (actual performance minus permutation
performance) are more important.
Left: models were trained on the full datasets, with different numbers of
samples available for each severity definition.
Right: models were trained on the intersection of samples with all labels
available for each definition. Note that Attributable and Pragmatic severity are
exactly the same for the intersection dataset.
OTU 120 (*Pseudomonas*) is not shown for IDSA severity on the full dataset nor
on the intersection dataset because it was removed during pre-processing due to
having near-zero variance.
**B)** Log~10~-transformed median relative abundances of the most important OTUs
on the full datasets, grouped by severity (shape). The vertical dashed line 
represents the limit of detection.
"}
![](figures/feature-importance.png)
:::
